---
title: "candy crush"
output: html_document
date: "2023-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# This sets the size of plots to a good default.
options(repr.plot.width = 5, repr.plot.height = 4)

# Loading in packages
library(readr)
library(dplyr)
library(ggplot2)
```
##The data set##
The dataset we will use contains one week of data from a sample of players who played Candy Crush back in 2014. The data is also from a single episode, that is, a set of 15 levels. It has the following columns:

player_id: a unique player id
dt: the date
level: the level number within the episode, from 1 to 15.
num_attempts: number of level attempts for the player on that level and date.
num_success: number of level attempts that resulted in a success/win for the player on that level and date.
The granularity of the dataset is player, date, and level. That is, there is a row for every player, day, and level recording the total number of attempts and how many of those resulted in a win.

```{r}
data <- read_csv('/Users/AmandaLin/Downloads/R-portfolio/candy_crush.csv')
#read.csv has a number of problems which the new read_csv function avoids 
head(data)
```
##Checking the data set##
```{r}
cat('number of players:',num_players) 

```
```{r}
print('period of data:') 
print(range(data$dt))
```
##Computing level difficulty##
###Bernoulli process###

```{r}
difficulty <- data %>%
  group_by(level) %>% 
  summarise(attempts = sum(num_attempts),wins = sum(num_success)) %>%
  mutate(p_win = wins/attempts)
head(difficulty)
```
##Plotting difficulty profile as line graph##
```{r}
ggplot(difficulty,aes(x=level,y=p_win))+
  geom_line()+
  scale_x_continuous(breaks = 1:15)+ #our data is from level 1 to level 15 
  scale_y_continuous(label = scales::percent) #makes the y-axis to be percentage
                       
```
##Spotting hard level##
###threashholdof difficulty = 10%###
```{r}
ggplot(difficulty, aes(x=level,y=p_win))+
  geom_line()+
  geom_point(color = 'blue')+
    geom_hline(yintercept = 0.1,linetype = 3,color='purple') #dashline is num3 in linetype
  scale_x_continuous(breaks = 1:15)+
  scale_y_continuous(label = scales::percent)
```

##Computing uncertainty##
###calculate standard error##
```{r}
difficulty <- difficulty %>%
  mutate(error = sqrt(p_win*(1-p_win)/attempts))
head(difficulty)
```
##Showing the uncertainty on the graph##

```{r}
ggplot(difficulty, aes(x=level, y=p_win))+
  geom_line()+
  geom_point(color = 'blue')+
  geom_errorbar(aes(ymin=p_win-error,ymax = p_win+error))+
  geom_hline(yintercept=0.1,linetype=3,color = 'purple')+
  scale_x_continuous(breaks = 1:15)+
  scale_y_continuous(label = scales::percent)
```
##Finding the possibility of a player complete the episode without losing a single time##
```{r}
p<-prod(difficulty$p_win) #prod function multiplies all the number in a vector together 
p
```

