---
title: "data scientist survey"
output: html_document
date: "2023-11-08"
---
```{r}
library(tidyverse)
library(ggpubr)
library(here)
library(skimr)
library(janitor)
library(lubridate)
library(ggrepel)
```
#load the data and print the first 10 rows
```{r}
responses <- read_csv("/Users/AmandaLin/Downloads/multipleChoiceResponses.csv")
head(responses, 10)
```
#extract the needed columns 

```{r}
data <- responses %>%
  select(WorkToolsSelect,LanguageRecommendationSelect,EmployerIndustry,WorkAlgorithmsSelect)
data <- data %>%   
  mutate(Respondent = row_number())%>%
  relocate(Respondent)
```
#focus on the tools and language that the survey respondents use at work 
#split the worktoolselect by comma


```{r}
tools <-data %>%
  mutate(work_tools = str_split(WorkToolsSelect,',')) %>%
  unnest(work_tools)

head(tools)
  
  
```
#most popular tools 
```{r}
tool_count <- tools %>%
  group_by(work_tools) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

head(tool_count)
```

#create a plot to show the most popular tools
```{r}
ggplot(tool_count, aes(x=fct_reorder(work_tools,count),y=count))+
  geom_bar(stat = 'identity')+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  xlab('Tools')
```

# Within the field of data science, there is a lot of debate among professionals about whether R or Python should reign supreme. You can see from our last figure that R and Python are the two most commonly used languages, but it's possible that many respondents use both R and Python. Let's take a look at how many people use R, Python, and both tools.
#create a new column called language preference

```{r}
debate_tools <- data %>%
  mutate(language_preference = 
           case_when(
    str_detect(WorkToolsSelect,'R') &! str_detect(WorkToolsSelect,'Python') ~'R',
        str_detect(WorkToolsSelect,'Python') &! str_detect(WorkToolsSelect,'R') ~'Python',
    str_detect(WorkToolsSelect,'R') & str_detect(WorkToolsSelect,'Python') ~'Both',
    TRUE ~ 'Neither'
  ))

head(debate_tools)
```

#plot the responses with python, r, or both
```{r}
debate_plot <- debate_tools %>%
  group_by(language_preference) %>%
  filter(language_preference!='Neither')%>%
  summarise(count=n())
ggplot(debate_plot,aes(x=fct_reorder(language_preference, count),y=count))+
  geom_bar(stat = 'identity')+
  xlab('R or Python')
```
#Language Recommendation#
#Do R or Python lover usually recommend their favorite language to new learners? 
```{r}
recommendations <- debate_tools %>%
  group_by(language_preference,LanguageRecommendationSelect) %>%
  summarise(count = n())%>%
  arrange(language_preference,desc(count))%>%
  mutate(row=row_number()) %>%
  filter(row <= 4)

recommendations



```

#Plot the recommendations to be a graph
```{r}
ggplot(recommendations, aes(x = LanguageRecommendationSelect, y=count))+
  geom_bar(stat='identity')+
  facet_wrap(~language_preference)
```

